Kelp morphometric properties and the link to hydrodynamic modelling

### 2018-09-10 chapter_1.Rmd
additional writing

#Abiotic drivers

@@

Depth does not aﬀect kelp ecosystems directly, however a change in depth often causes ﬂuctuations or changes in other environmental variables such as water motion, light and temperature. Water motion also decreases with depth, and some kelps better suited to deeper environments (*L. pallida*) replace those in the shallows (*E. maxima*) (Dayton 1985; Gerard 1982). The increase in depth can lead to a decrease in sunlight penetration, with some species better adapted for low-light conditions than others, such as (*L. pallida*). Temperature may also change along a depth gradient due to a reduction in sunlight penetration. (Dayton 1985; Gerard 1982). Therefore depth does not directly play a role in kelp functioning but may alter more inﬂuential factors such as light and water motion.

@@

Other than temperature, wave exposure is also recognised as an important driver of the marine environment. Like temperature, wave exposure can act on regional scales and has been shown to play a role in determing distribution, abundance, diversity, composition and productivity of benthic and rocky shore communties. For example, increasing degrees of exposure may positively influence the amount of area available to trap light on macroalgal fronds, as well as increasing nutrient uptake through increased turbulence in the boundry layer [@lobban1994]. @lewis found that degrees of wave exposure vertical structure rocky shore communities...**work in progress**...

>>>>

#Kelp environmental drivers

The important environmental drivers of kelp idividuals and communities include light, substrata, salinity, sedimentation, nitrients, temperature and water motion. Although studies have investigated the effects of important environmental drivers, the roles these factors play is often difficult to evaluate as such factors may never be fully independent of each other, i.e. environmental factors are to some extent temperature dependent. Multifactorial studies have attempt to explain combined affects, however these studies are often limited to investigating combination of two or three environmental drivers as inclusion of too many factors can lead to results that are difficult to interept. Environmetal factors are highly variable on temporal and spatial scales, and their effects may also dependent on life-stage adding a further layer of complexity to investigations.

@@

Depth does not aﬀect kelp ecosystems directly, however a change in depth often causes ﬂuctuations or changes in other environmental variables such as water motion, light and temperature. Water motion also decreases with depth, and some kelps better suited to deeper environments (*L. pallida*) replace those in the shallows (*E. maxima*) [@Dayton1985; @Gerard1982]. The increase in depth can lead to a decrease in sunlight penetration, with some species better adapted for low-light conditions than others, such as (*L. pallida*). Temperature may also change along a depth gradient due to a reduction in sunlight penetration [@Dayton1985; @Gerard1982]. Therefore depth does not directly play a role in kelp functioning but may alter more inﬂuential factors such as light and water motion.

@@

Other than temperature, wave exposure is also recognised as an important driver of the marine environment. Like temperature, wave exposure can act on regional scales and has been shown to play a role in determing distribution, abundance, diversity, composition, growth [@Cousens1982] and productivity of macroalgae communties. The width, vertical zonation and diversity of algal communities often change predictably along gradients of exposure. This is also true for variation in morphological characters of individuals may also vary with wave exposure, such as frond width and degree of 'branching' [@moss1948]. @moss1948 investigated the anatomy, chemical composistion of *Fucus spiralis* at three sites that varied in wave exposure (sheltered, medium exposure and exposed). The authors found that individuals in exposed sites showed less branching of thalli as well as variation in organic nitrogen, mannitol, laminarian and alginic acid in relation to wave exposure between study sites.

For example, increasing degrees of exposure may positively influence the amount of area available to trap light on macroalgal fronds, as well as increasing nutrient uptake through increased turbulence in the boundry layer [@lobban1994]. @lewis found that degrees of wave exposure vertical structure rocky shore communities...**work in progress**...

### 2018-09-10 chapter_1.tex
additional writing

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Kelp morphometric properties and its link to the hydrodynamic modelling},
            pdfauthor={Ross Coppin},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

  \title{Kelp morphometric properties and its link to the hydrodynamic modelling}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Ross Coppin}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{26 February 2017}

@@

Seaweeds, browns in particular, are capable of adapting morphological
characteristics to persist in changing and variable ocean environments.
Changes in morphology have been shown to increase photosynthetic
ability, enhance nutrient uptake, and reduce drag with regards to
changes in temperature and wave exposure. Wave exposure has been shown
to be an important driver of seaweed morphology, as the main mechanism
of seaweed mortality is through the mechanism of dislodgment. Changing
morphology, such as decreasing frond area or strengthening through stipe
circumference increase/reduces drag and increases the probability of
survival. However, locally adapted seaweed may still be dislodged in
pulse disturbance events such as storms, and may raft far distances and
wash up on beaches. An example would be \emph{Ecklonia maxima} which is
a conspicuous brown seaweed along the coastline and beaches around South
Africa. Therefore, the beach-cast may not always originate from adjacent
kelp populations but rather from other regions which the individual is
adapted for. Therefore, the kelp morphology may act as a proxy for
investigating coastal currents and changes thereof. The advances in
ocean hydrodynamic modelling has made great progress and has been
applied in a variety of ways. Therefore, this study will use advances in
hydrodynamic modelling in combination with kelp morphological
characteristics to investigate coastal currents. Once the model has been
established it may be applied in other ways, such as investigating the
transport of micorplastics along the South African coastline. The harms
of microplastics to the marine environment has gained much traction in
recent years, but research in South Africa is lacking. The coupling of
kelp morphology and the ability to simulate hydrodynamic processes can
greatly improve our understanding of transport pathways and likely
locations of accumulation. This in turn may inform management decisions
with regards to eliminating and managing marine pollution in South
Africa.

@@

In recent years there has been growing attention on microplastics and
the deterimental role they play in the enviornment, particulary the
ocean. Microplastics are tiny plastic granuales used as scrubbers in
cosmetics and air-blasting, and small plastic fragments that originate
from larger pieces of plastic known as macroplastics. The potential
harms of small pieces of plastic was highlighted in the 1970's and
renewed interest has lead to research showing that microplastics in the
ocean are widespread. The small size of microplastics makes them
bio-available to biota in the food-web which may cause problems with an
organisms physiological functioning. Furthermore, the relatively large
surface area and composition of microplastics provides an environmeht
that is able of adhering to organic pollutants. In other words
microplastics also act as a vector for transport and assimilation of
organic polluants. Therefore, microplastics may also aid in the
assimilation of organic pollutans into all levels of the food web.

@@

\section{Kelp morphology drivers}\label{kelp-morphology-drivers}

@@

Light is an important factor for kelp survival, however if light is
limited or excessive this may negatively impact kelp survival or growth.
Much of the past research into the role light plays into the functioning
of kelp {[}@Bruhn1996);@Belseth2012{]}. For instance, solar ultraviolate
radiation has been shown to aﬀect sub-canopy Ecklonia radiata
sporophytes when the canopy of mature \emph{Ecklonia radiata} was
removed {[}@Wood 1987{]}. The sub-canopy sporophytes experienced tissue
damage,photopigment destruction,reducedgrowthand decreased survivorship,
thus inhibiting their settlement and survival {[}@Wood 1987{]}.
Laboratory experiments revealed that the UV component of radiation,
rather than intense radiation itself, was responsible for the eﬀects
mentioned above. High light stress has negative eﬀects, such as
photoinhibition and photo-damage on \emph{Ecklonia cava} sporophytes
{[}@Altamirano2004{]}. @Altamirano2004 found that \emph{Ecklonia cava}
is more vulnerable to light stress conditions, and less likely to
recover under unfavourable conditions {[}@Altamirano2004{]}. @Bolton1985
showed that under sub-saturating irradiances and supra- optimal
temperatures \emph{Ecklonia maxima} to showed a decrease in reproductive
rates and an increase in cell production. A interesting ﬁnding of this
study was that despite the decrease in reproductive rates, the ﬁnal egge
production per female was greater under these conditons. The authors
interpreted this an ecological adaption that may increase survival rates
under times of stress or non - ideal conditions {[}@Bolton1985{]}. Light
is also important for survival and functioning of benthic algal species;
a study by @Toohey2004 investiagted the possible link of light and
thallus scour of the kelp \emph{Ecklonia radiata canopy} on the
associated foliose algal assemblage. The authors concluded that a
reduction in light, rather than thallus scour has an eﬀect on foliose
algal assemblage. Reduced light was associated with reduced diveristy
and an marked increase abundance in certian foliose species, which were
able to acclimiate more light under low irradiances compared to other
species {[}@Toohey2004{]}. Kelp may also compete intra-specﬁcally or
inter-speciﬁcally for light. For instance, @Clark2004 investiagted eﬀect
of canopy shading of \emph{Macrocystis pyrifera} and understorey canopy
shading of \emph{Pterygophora californica} individually and in
combination of benthic macro-algae assemblage. Under all treatments
there was a sigiﬁance change in macro-algal assemblage, with certian
species able to recover in abundance more readily than others. The
authors devided these species into two categories; ``shade-tolerant''
and ``light adapted'', and conlcuded that certian foliose species are
better adapted for diﬀerent light conidtions in accordance with canopy
cover. In South Africa, the eﬀect of \emph{Ecklonia maxima} harvesting
on sub-story kelp, kelp population structure and kelp recruitement was
investigated by @Rothman2006. The results showed no signiﬁcant diﬀerence
in the three factors mentioned earlier and @Rothman2006 concluded that
factors other than light play a role in structuring kelp communities.

@@

Nutrients
Theimportanceofnutrientsinthefunctioningofkelpsiswellunderstood
{[}@Dayton1985; @Gaylord2012{]}. Dissolved nitrogen, and in particular
nitrate, are important; however research has also placed emphasis on
phosphate and other trace compounds for functioning of kelps
{[}@Dayton1985{]}. Additionally, some kelps have the ability to store
inorganic nitrogen in order to compensate for periods of low nutrient
availability, which has been observed for Laminaria and Macrocystis
{[}@Dayton1985; @Gaylord2012{]}. Nutrient stratiﬁcation is also an
important factor, particularly for canopy type kelps. The concentration
of nutrients at the surface is important to the functioning and
maintenance of the canopy. For instance kelp canopies in California
often deteriorate in the summer months when surface nitrate levels are
low {[}@Jackson1977{]}. Water motion is important in the assimilation of
nutrients from the water column, and kelps have been shown to adapt
blade morphology in order to create more turbulence around the boundary
layer of the frond to enhance nutrient assimilation (Wheeler 1980).
Temperature has also been closely linked with nutrient concentrations.
Nutrients are often in higher concentrations in the water column during
low temperature events. This is often an indication of an ``up-welling''
event, which brings cold and nutrient rich waters from the bottom to the
surface of the water column. However temperature can play a direct role
in the uptake of nutrients through eﬀects on algal metabolism; however
this may vary from species to species {[}@Raven1988{]}. However most
studies investigate multiple characteristics in one study
(i.e.~temperature and light and water motion), while studies
investigating the direct role of temperature in nutrient assimilation is
lacking in marine algae is lacking.

Depth does not aﬀect kelp ecosystems directly, however a change in depth
often causes ﬂuctuations or changes in other environmental variables
such as water motion, light and temperature. Water motion also decreases
with depth, and some kelps better suited to deeper environments
(\emph{L. pallida}) replace those in the shallows (\emph{E. maxima})
(Dayton 1985; Gerard 1982). The increase in depth can lead to a decrease
in sunlight penetration, with some species better adapted for low-light
conditions than others, such as (\emph{L. pallida}). Temperature may
also change along a depth gradient due to a reduction in sunlight
penetration. (Dayton 1985; Gerard 1982). Therefore depth does not
directly play a role in kelp functioning but may alter more inﬂuential
factors such as light and water motion.

\section{Sedimentation}\label{sedimentation}

Sedimentation and scour can be highly destructive to kelp forests with
most detrimental impacts on the development stages such as spores,
gametophytes and young plants (Dayton 1985), and has been linked to
seasonal trends relating to wind direction {[}@Moore1972{]}.
@Devinny1978 found that sedimentation greatly reduced spore settlement
and which in turn greatly reduces the probability of survival for
Macrocystis Pyrifera gametophytes. Some species seem more resistant to
the eﬀects of sedimentation such as Laminaria farlowii and Eisena aborea
compared to others such as Macrocystis Pyrifera (Dayton 1985). The
impacts of sedimentation is not limited to the sub-tidal but has also
been shown to eﬀect inter-tidal kelp as well. For example a study by
{[}@Schiel2006{]} investigated the eﬀects of sedimentation on the
post-settlement stages of habitat-forming seaweeds. The results showed
high mortality of germlings at sites that had high sedimentation, and
was further supported through laboratory tests. Grazing was found to be
negligible and mortality of gametophytes was attributed to
sedimentation, although its eﬀects did vary with molluscan grazing
{[}@Schiel2006{]}. The presence of kelps themselves can also eﬀect the
degree of sedimentation. For instance, @Wernberg2005 investigated the
ability of Ecklonia radiata (canopy forming species) to modify the
physical environment and the related eﬀects on associated foliose algae.
An increase in kelp densities was marked by a signiﬁcant reduction of
light and sedimentation, and found that sedimentation did not have a
signiﬁcant eﬀect on associated foliose algae {[}@Wernberg2005{]}.
Therefore it is often diﬃcult to separate the direct eﬀects of each of
the environmental variables as they are often related to one another.
However, sea surface temperatures is one of the main predictors of
marine biological systems, as it has been shown to relate to
physiological and ecological processes at all levels {[}@Andersen2013;
@Bolton1985; @Bolton2012; @Dayton1985{]}. Kelp forests are no exception,
with temperature being the main limiting factor {[}@Dayton 1985{]}.

@@

Temperature Biogeographical shifts of kelp species can already be seen
in areas where ocean warming is occurring. In Western Australia an
increase in temperature has caused a biogeographical shift in kelp and
urchins, which in turn has caused many kelp forests to shift to the less
productive algal turfs {[}@Kelly2005{]};a review of the literature shows
a consensus that biogeographical distributions of marine species are
largely temperature related {[}@Dayton1985; @Jeﬀree1994; @Schils2006;
@Sorte2010; @Bolton2012{]}. Other examples include Norway and eastern
North America where ocean warming has facilitated kelp die-oﬀ and an
increase in herbivore grazing rates {[}@Dayton1985{]}. The increase in
grazing rates and the colonization of other algal species into these
areas make it almost impossible for kelps to recover {[}@Kelly2005{]}.
The majority of kelp species are artic and temperate organisms, and the
warming of ocean temperatures is expected to cause a poleward
biogeographical shift of species {[}@Bolton2012{]}. There is evidence to
suggest that South African kelp forests are expanding due to ocean
cooling {[}@Bolton2012{]}, possibly driven by an intensiﬁcation and
increase in coastal upwelling {[}@Blamey2012, @Blamey2015{]}. In South
Africa there has been a biogeographical shift eastward along the coast
due to a change in inshore temperature regime, making South Africa no
exception to changing ocean temperatures {[}@Bolton2012{]}. Macroalgae,
such as kelps, can react to an increase in surface temperatures in one
of three ways: they can migrate, adapt and die {[}@Biskup2014{]}. A
study by @Biskup2014 investigated the functional response of two kelp
species (\emph{Laminaria ochroleuca} and \emph{Saccorhiza polyschides})
to rising sea temperatures. The functional responses of Saccorhiza
polyschides was measured for both the subtidal and intertidal habitats,
to see what aﬀect non- optimal conditions (intertidal zone) had on the
kelps {[}@Rinde2005{]}. The study found that Laminaria ochroleuca
exhibited a poor ability to acclimatise and was dependent on the kelp's
life history traits {[}@Biskup2014{]}. Therefore annual kelp species are
more likely to survive under non-ideal condition, and the intertidal
Saccorhiza polyschides, compared to the subtidal, showed a higher
physiological ﬂexibility to changing conditions {[}@Biskup2014{]}. This
may be because the intertidal zone undergoes far more change than the
subtidal and therefore kelps in the intertidal are forced to adapt to
harsher conditions where ﬂuctuations in temperature, sunlight, turbidity
and water motion are common. The eﬀects on temperature have also been
investigated by @Wernberg2010. The study looked at resilience of kelp
beds along a latitudinal temperature gradient. Kelp abundance is likely
to decline with the predicted warming of ocean waters @Wernberg2010 and
although kelps have the ability to acclimatize and adjust their
metabolic performance, which in turn allows them to change their
physiological performance to mitigate the seasonal ﬂuctuations in
temperature, this acclimatization is done at a cost @Wernberg2010. It is
therefore important to understand the ecological implications of this
cost in order to better understand the future impacts of rising sea
temperatures @Wernberg2010. Results from the study showed that kelp beds
in the northern regions had a lower resilience to perturbations, such as
physical and biological disturbance, disease, competition from invasive
species and decreasing water quality. This was as a result of low
abundance of recruits, reduced physiological responsiveness and reliance
on mature adult kelp for net recruitment, growth of recruits and
recovery of canopy loss @Wernberg2010. Even though @Wernberg2010
acknowledged that it was diﬃcult to exclude the inﬂuences of other
environmental parameters (sunlight, nutrients and wave exposure),
temperature was most likely the main environmental driver of these
eﬀects. While kelps may be able to maintain populations in the warmer
upper limits of their distributions, the cost of acclimation will make
them more susceptible to other stressors such as disturbance from
anthropogenic sources, disease, competition and decreasing water
quality. Changes in temperature can also impact the distribution of
kelps.

>>>>

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{eurosym}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Kelp morphometric properties and the link to hydrodynamic modelling},
            pdfauthor={Ross Coppin},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

  \title{Kelp morphometric properties and the link to hydrodynamic modelling}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Ross Coppin}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{26 February 2017}

@@

Seaweeds, browns in particular, are capable of adapting morphological
characteristics to persist in changing and variable ocean environments.
Changes in morphology have been shown to increase photosynthetic
ability, enhance nutrient uptake, and reduce drag with regards to
changes in temperature and wave exposure. Wave exposure has been shown
to be an important driver of seaweed morphology, as the main mechanism
of seaweed mortality is through the dislodgment. Changing morphology
reduces drag and increases the probability of survival. However, locally
adapted seaweed may still be dislodged in pulse disturbance events such
as storms, and may raft far distances and wash up on beaches. An example
would be \emph{Ecklonia maxima} which is a conspicuous brown seaweed
along the coastline and beaches around South Africa. Therefore, the
beach-cast may not always originate from adjacent kelp populations but
rather from other regions which the individual is adapted for.
Therefore, the kelp morphology may act as a proxy for investigating
coastal currents and changes thereof. The advances in ocean hydrodynamic
modelling has made great progress and has been applied in a variety of
ways. Therefore, this study will use advances in hydrodynamic modelling
in combination with kelp morphological characteristics to investigate
coastal currents. Once the model has been established it may be applied
in other ways, such as investigating the transport of micorplastics
along the South African coastline. The harms of microplastics to the
marine environment has gained much traction in recent years, but
research in South Africa is lacking. The coupling of kelp morphology and
the ability to simulate hydrodynamic processes can greatly improve our
understanding of transport pathways and likely locations of
accumulation. This in turn may inform management decisions with regards
to eliminating and managing marine pollution in South Africa.

@@

In recent years there has been growing attention on plastic pollution,
particulary in the ocean. Plastic pollution can be in the form of macro-
and microplastics. Microplastics are tiny plastic granuales used as
scrubbers in cosmetics and air-blasting, and small plastic fragments
that originate from larger pieces of plastic known as macroplastics,
while macroplastics\ldots{}\textbf{insert definition here}\ldots{} The
potential harms of of plastic pollution in the marine environment was
highlighted in the 1970's and renewed interest has lead to research
showing that plastic pollution in the ocean are widespread. Plastics may
become bio-available to biota in the food-web which may cause problems
with an organisms physiological functioning. Furthermore, the relatively
large surface area and composition of microplastics provides an
environmeht that is able of adhering to organic pollutants. In other
words microplastics also act as a vector for transport and assimilation
of organic polluants.

@@

\section{Kelp environmental drivers}\label{kelp-environmental-drivers}

The important environmental drivers of kelp idividuals and communities
include light, substrata, salinity, sedimentation, nitrients,
temperature and water motion. Although studies have investigated the
effects of important environmental drivers, the roles these factors play
is often difficult to evaluate as such factors may never be fully
independent of each other, i.e.~environmental factors are to some extent
temperature dependent. Multifactorial studies have attempt to explain
combined affects, however these studies are often limited to
investigating combination of two or three environmental drivers as
inclusion of too many factors can lead to results that are difficult to
interept. Environmetal factors are highly variable on temporal and
spatial scales, and their effects may also dependent on life-stage
adding a further layer of complexity to investigations.

@@

Light is an important factor for kelp survival, however if light is
limited or excessive this may negatively impact kelp survival or growth.
Much of the past research into the role light plays into the functioning
of kelp {[}@Bruhn1996);@Belseth2012{]}. For instance, solar ultraviolate
radiation has been shown to aï¬\euro{}ect sub-canopy Ecklonia radiata
sporophytes when the canopy of mature \emph{Ecklonia radiata} was
removed {[}@Wood 1987{]}. The sub-canopy sporophytes experienced tissue
damage,photopigment destruction,reduced growthand decreased
survivorship, thus inhibiting their settlement and survival {[}@Wood
1987{]}. Laboratory experiments revealed that the UV component of
radiation, rather than intense radiation itself, was responsible for the
eï¬\euro{}ects mentioned above. High light stress has negative
eï¬\euro{}ects, such as photoinhibition and photo-damage on
\emph{Ecklonia cava} sporophytes {[}@Altamirano2004{]}. @Altamirano2004
found that \emph{Ecklonia cava} is more vulnerable to light stress
conditions, and less likely to recover under unfavourable conditions
{[}@Altamirano2004{]}. @Bolton1985 showed that under sub-saturating
irradiances and supra- optimal temperatures \emph{Ecklonia maxima} to
showed a decrease in reproductive rates and an increase in cell
production. An additional ï¬nding of this study was that despite the
decrease in reproductive rates, the ï¬nal egge production per female
was greater under these conditons. The authors interpreted this an
ecological adaption that may increase survival rates under times of
stress or non - ideal conditions {[}@Bolton1985{]}.

Depth does not aï¬\euro{}ect kelp ecosystems directly, however a change
in depth often causes ï¬‚uctuations or changes in other environmental
variables such as water motion, light and temperature. Water motion also
decreases with depth, and some kelps better suited to deeper
environments (\emph{L. pallida}) replace those in the shallows (\emph{E.
maxima}) (Dayton 1985; Gerard 1982). The increase in depth can lead to a
decrease in sunlight penetration, with some species better adapted for
low-light conditions than others, such as (\emph{L. pallida}).
Temperature may also change along a depth gradient due to a reduction in
sunlight penetration. (Dayton 1985; Gerard 1982). Therefore depth does
not directly play a role in kelp functioning but may alter more
inï¬‚uential factors such as light and water motion.

@@

The importance of nutrients in the functioning of kelps is well
understood {[}@Dayton1985; @Gaylord2012{]}. Dissolved nitrogen, and in
particular nitrate, are important; however research has also placed
emphasis on phosphate and other trace compounds for functioning of kelps
{[}@Dayton1985{]}. Additionally, some kelps have the ability to store
inorganic nitrogen in order to compensate for periods of low nutrient
availability, which has been observed for Laminaria and Macrocystis
{[}@Dayton1985; @Gaylord2012{]}. Nutrient stratiï¬cation is also an
important factor, particularly for canopy type kelps. The concentration
of nutrients at the surface is important to the functioning and
maintenance of the canopy. For instance kelp canopies in California
often deteriorate in the summer months when surface nitrate levels are
low {[}@Jackson1977{]}. Water motion is important in the assimilation of
nutrients from the water column, and kelps have been shown to adapt
blade morphology in order to create more turbulence around the boundary
layer of the frond to enhance nutrient assimilation (Wheeler 1980).
Temperature has also been closely linked with nutrient concentrations.
Nutrients are often in higher concentrations in the water column during
low temperature events. This is often an indication of an
â\euro{}œup-wellingâ\euro{} event, which brings cold and nutrient rich
waters from the bottom to the surface of the water column. Temperature
can play a direct role in the uptake of nutrients through eï¬\euro{}ects
on algal metabolism; however this may vary from species to species
{[}@Raven1988{]}.

@@

Temperature is a driver of kelp species distributions and
ecophysiological processes, as well as a lesser role in morphological
adaption\ldots{}\textbf{example here}\ldots{}The majority of kelp
species are artic and temperate organisms, and the warming of ocean
temperatures is expected to cause a poleward biogeographical shift of
species {[}@Bolton2012{]}. There is evidence to suggest that South
African kelp forests are expanding due to ocean cooling
{[}@Bolton2012{]}, possibly driven by an intensiï¬cation and increase
in coastal upwelling {[}@Blamey2012, @Blamey2015{]}. In South Africa
there has been a biogeographical shift eastward along the coast due to a
change in inshore temperature regime, making South Africa no exception
to changing ocean temperatures {[}@Bolton2012{]}. Macroalgae, such as
kelps, can react to an increase in surface temperatures in one of three
ways: they can migrate, adapt and die {[}@Biskup2014{]}. A study by
@Biskup2014 investigated the functional response of two kelp species
(\emph{Laminaria ochroleuca} and \emph{Saccorhiza polyschides}) to
rising sea temperatures. The functional responses of Saccorhiza
polyschides was measured for both the subtidal and intertidal habitats,
to see what aï¬\euro{}ect non- optimal conditions (intertidal zone) had
on the kelps {[}@Rinde2005{]}. The study found that Laminaria ochroleuca
exhibited a poor ability to acclimatise and was dependent on the
kelpâ\euro{}™s life history traits {[}@Biskup2014{]}. Therefore annual
kelp species are more likely to survive under non-ideal condition, and
the intertidal Saccorhiza polyschides, compared to the subtidal, showed
a higher physiological ï¬‚exibility to changing conditions
{[}@Biskup2014{]}. This may be because the intertidal zone undergoes far
more change than the subtidal and therefore kelps in the intertidal are
forced to adapt to harsher conditions where ï¬‚uctuations in
temperature, sunlight, turbidity and water motion are common. The
eï¬\euro{}ects on temperature have also been investigated by
@Wernberg2010. The study looked at resilience of kelp beds along a
latitudinal temperature gradient. Kelp abundance is likely to decline
with the predicted warming of ocean waters @Wernberg2010 and although
kelps have the ability to acclimatize and adjust their metabolic
performance, which in turn allows them to change their physiological
performance to mitigate the seasonal ï¬‚uctuations in temperature, this
acclimatization is done at a cost @Wernberg2010\ldots{}\textbf{link to
paragraph on kelp morphology}\ldots{}

\section{Wave exposure}\label{wave-exposure}

Other than temperature, wave exposure is also recognised as an important
driver of the marine environment. Like temperature, wave exposure can
act on regional scales and has been shown to play a role in determing
distribution, abundance, diversity, composition and productivity of
benthic and rocky shore communties. For example, increasing degrees of
exposure may positively influence the amount of area available to trap
light on macroalgal fronds, as well as increasing nutrient uptake
through increased turbulence in the boundry layer {[}@lobban1994{]}.
@lewis found that degrees of wave exposure vertical structure rocky
shore communities\ldots{}\textbf{work in progress}\ldots{}

\section{Hydrodynamic modelling}\label{hydrodynamic-modelling}

Traditional ecological measures of wave exposure usually incorportates
integrative measures of hyrdodynamic conditions at a particular site.
More specifically, it is the integration of mechanical processes and the
influence that is has on the ecology of nearshore communities.

Wave exposure may be modelled through various methods which range from
simple cartographic to more advanced numerical wave models.
Cartographical models can be qualitative or quantitative and were
designed for the need of wave exposure measures to explain ecological
distributions. A simple set of calculations on coastline and wind data,
and relatively small input data sets are required. These are regarded as
``fetch-based models'', which measure the length of open water
associated with a site along a straight line. The output of such an
approach is a simplified estimate of the potential wave energy for a
specific set of sites. Advances in cartographical methods using
fetch-based models has allowed for wave exposure measurments for larger
areas, and has been suggested as a method for predicting macroalgal
community structure {[}@burrows2008{]}. An example of such a model is
the ``BioEx model'' which was developed by @baardseth1970 to estimate
wave exposure over large regions. BioEx requires frequency, strength and
direction of winds, weighted by degree of exposure within various
directions. BioEx is calculated as the sum of the index developed at
different spatial scales (local, fjord and open). This method has been
used in mapping of marine coastal biodiveristy {[}@Rinde2004{]}.

@lindegarth2005 critized this approach, arguing that the choice of wave
exposure method can influence ecological inference. The authors also
highlighted the need for objective, reproducible and quantitative
studies comparing exposure indices {[}@lindegarth2005{]}. A study by
@sundbald\ldots{}\textbf{include description of study here}\ldots{}Other
authors, such as @hill2010, have argued that these simple measures can
be improved upon by including bathymetery data which allows the
incorporation of diffraction into the calculation. Diffraction is
topographically induced variations in wave direction. A model
incorporating this complexity was developed by @isaeus2004, and is known
as the ``simplified wave model'' (SWM). The model uses measurements of
wind strength, fetch and empirically derived algorithims to mimic
diffraction.

Advances in numerical modelling have been founded on physical wave
theory on how a wave ``behaves''. This approach is based on a
theoretical persepctive rather than the need to answer ecological
questions. Besides diffraction, these models incorporate more complexity
by including wind forcing, wave-to-wave interactions and loss of energy
due to friction and wave breaking. Numerical models have a variety of
applications and are often incorporated within hydrodynamic general
circulation models and are used operationally for forecasting the sea
state {[}@hasselmann1988; @booij1999; @smith2001{]}. The downside of
advanced numerical models is that are computationally intensive which
creates limitations for large scale simulations. Therefore, their
application along long stretches of variable coastline, inshore
environments and ocean-wide simulations limited due to the poor spatial
coverage. However, this models can be designed for local or site
specific coverage, provided the correct data is available.
